A Decision Procedure for String Constraints with String-Integer Conversion and Flat Regular Constraints

### 2021-05-20 FMCAD2021/6_impl_exp.tex
section VI

\paragraph{The STRINGHASH benchmark suite} 
This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ such that for each $w= a_1 \ldots a_n$, 
${\sf hash}(w) = \sum \limits_{i = 1}^n a_i p^{n-i} \bmod m$, where $p, m \in \Int^+$ with $p, m \ge 2$. It is easy to see that ${\sf hash}(w)$ can be seen as a generalization of $\parseInt$ followed by a modulo operation. In particular, if $\Sigma = \Sigma_{\sf num}$ and $p = 10$, then ${\sf hash}(w) = \parseInt(w) \bmod m$. 
Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, e.g. CVC4, Z3, and Trau, we assume $\Sigma =\Sigma_{\sf num}$ and $p = 10$, so that the string hash function can be encoded by $\parseInt$.

The string constraints in the STRINGHASH benchmark suite are of the form 
$$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$$ 
where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
The six groups of string constraints are characterized by different classes of regular constraints, namely, $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, $w^+_1 w^+_2 6789$, $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, and $\Sigma_{\sf num}^* 6789$, where $w_1,w_2 \in \Sigma^+_{\sf num}$. The first three are flat languages, while the latter three are not. Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number). We generate the SMTLib2 files for these string constraints, as inputs to the SMT solvers. For the {\paexp}-solver, we do the following:   
\begin{itemize}
\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
%
\item For non-flat regular constraints, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
$\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
%
%the first part is a single digit and the second part is a string length $l$, which can be seen as a FL with 2 cycles and force the second cycle do not repeat. 
%
We iterate the following procedure until a model is found or the time limit is reached: Initially, set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$. For each assignment, we encode the resulting string constraint into an {\paexp} formula with only one exponential variable. If the resulting {\paexp} formula is unsatisfiable, then we increase $k$ by 1 and repeat this process. 
\end{itemize}
We would like to remark that the flattening strategy for non-flat regular constraints is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns. On the other hand, we consider patterns of the form e.g. $(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode.

>>>>

\paragraph{The STRINGHASH benchmark suite} 
This suite comprises six groups of string constraints generated from the string hash functions ${\sf hash}(w)$ encoded by $\parseInt$.
%Aiming at a comparison with the state-of-the-art string constraint solvers that support $\parseInt$, we assume that $\Sigma =\Sigma_{\sf num}$ and $p = 10$ and the string hash function is encoded by $\parseInt$.
%, so that the string hash function can be encoded by $\parseInt$.

The string constraints in the STRINGHASH benchmark suite are of the form 
$$\svarx \in \Aut \wedge \left( \parseInt(\svarx) \bmod m \right) \bmod m^\prime = 0   \notag \wedge \len(\svarx) < 100,$$ 
where $\Aut$ is an FA, $m, m^\prime \ge 2.$ 
The six groups of string constraints are characterized by different classes of regular constraints, namely, $12345w^+_1 w^+_2$,  $12345 w^+_1  w^+_2 6789$, $w^+_1 w^+_2 6789$, $12345\Sigma_{\sf num}^*$, $12345\Sigma_{\sf num}^* 6789$, and $\Sigma_{\sf num}^* 6789$, where $w_1,w_2 \in \Sigma^+_{\sf num}$. The first three are flat languages, while the latter three are not. Each group has 100 randomly generated problem instances, where $m$ is a randomly chosen prime number in the interval $[10^2, 10^5]$ and $1 \le m' < m$ ($m'$ is not necessarily a prime number). We generate the SMTLib2 files for these string constraints, as inputs to the string constraint solvers. On the other hand, for the {\paexp}-solver, we do the following:   
\begin{itemize}
\item For flat regular constraints, we generate the {\paexp} formulas corresponding to the string constraints, as the inputs to the {\paexp}-solver.
%
\item For non-flat regular constraints, we use flat languages $(a)^* (b_{1} \ldots b_{k})$ to under approximate 
$\Sigma_{\sf num}^*$,  where $a, b_1,\ldots, b_k \in \Sigma_{\sf num}$. 
%
%the first part is a single digit and the second part is a string length $l$, which can be seen as a FL with 2 cycles and force the second cycle do not repeat. 
%
We iterate the following procedure until a model is found or the time limit is reached: Initially, set $k=1$ and iterate by assigning $0, \ldots, 9$ to $a$. For each assignment, we encode the resulting string constraint into an {\paexp} formula with only one exponential variable. If the resulting {\paexp} formula is unsatisfiable, then we increase $k$ by 1 and repeat this process. 
\end{itemize}
We would like to remark that the flattening strategy for non-flat regular constraints here is a strict generalization of that in \cite{Parosh:20:PLDI}: Patterns of the form $0^*(b_1...b_k)$ were considered therein and {\pa} formulas are sufficient to encode such patterns. On the other hand, we consider patterns of the form e.g. $(a)^* (b_1 \ldots b_k)$ (where $a \in \Sigma_{\sf num}$ can be nonzero), which requires {\paexp} formulas to encode in general.
